Analyzing Arrest Patterns and Racial Representation in Toronto

### 2024-09-24 paper/paper.qmd
Add files via upload

---
title: "Analyzing Arrest Patterns and Racial Representation in Toronto"
author: "Akshat Aneja"
abstract: The city of Toronto has an extremely diverse population in terms of its ethnicity and culture. This paper examines arrest data from the Toronto police service between 2020 and 2021. It focuses on the distribution of arrests by ethnic identity, age and booking status. This is done to check whether policing practices exhibit systemic bias. By cleaning and analyzing the records, trends are explored over major ethnic groups. Despite concerns over racial bias, the data suggests that arrests are influenced by objective factors, though there still may be disparities due to socio-economic status. These findings may help in providing insights that may enhance transparency in policing practices.
date: today
date-format: long
number-sections: true
format: pdf
toc: true
thanks: "Code and data supporting this analysis is available at: https://github.com/Akshat211202/Analysis-of-bias-in-arrests/"
bibliography: references.bib
---

@@

In recent years, debates surrounding racial bias in policing have intensified, particularly regarding its disproportionate impact on marginalized communities. This issue is especially pertinent in Toronto, a culturally and ethnically diverse city. Discussions on police reform often reflect historical connections to slavery and contemporary racial disparities, raising concerns about whether law enforcement treats marginalized communities differently, potentially leading to racial profiling.

A report from Toronto Police Services indicates that Black, East/Southeast Asian, Middle Eastern, and Latino individuals are overrepresented in use-of-force incidents relative to their population in enforcement actions. Despite efforts by Toronto Police to promote a more equitable society, these findings suggest the continued presence of racial bias in arrests.This paper analyzes arrest data from 2020 and 2021 to explore interactions between police and various ethnic groups. Disparities in arrest rates may reflect broader socio-economic factors such as poverty and unemployment, which disproportionately affect marginalized communities and contribute to higher involvement in criminal activities.

Arrest data provides a quantifiable view of law enforcement actions and their correlation with ethnicity and age. [@sec-data], outlines the data cleaning process using R, followed by an analysis of arrest distributions across age groups to identify potential trends. Age is a critical factor in this analysis, as it may reveal patterns of concern. [@sec-discussion], presents findings and discusses their implications. The study seeks to contribute to a deeper understanding of policing in Toronto, providing insights for policy reforms and community engagement efforts

>>>>

---
title: "Analyzing Arrest Patterns and Racial Representation in Toronto"
author: "Akshat Aneja"
abstract: Toronto is known for its ethnic and cultural diversity, making it a relevant setting for examining potential biases in law enforcement. This paper analyzes arrest data from the Toronto Police Service for the years 2020 and 2021, focusing on the distribution of arrests by ethnic identity, age, and booking status to assess whether systemic bias exists in policing practices. The analysis reveals that, while arrests appear to be driven by objective factors, disparities related to socio-economic status may still exist. These findings provide important insights that can help improve transparency and accountability in policing, contributing to more equitable law enforcement practices.
date: today
date-format: long
number-sections: true
format: pdf
toc: true
thanks: "Code and data supporting this analysis is available at: https://github.com/Akshat211202/Analysis-of-bias-in-arrests/"
bibliography: references.bib
---

@@

Racial bias in policing has become a central issue in public discourse, particularly regarding its disproportionate impact on marginalized communities. This issue is of special significance in cities like Toronto, known for its cultural and ethnic diversity. Amid calls for police reform, historical ties to slavery and contemporary racial disparities often dominate discussions, prompting concerns that marginalized groups may experience discriminatory treatment, including racial profiling, in interactions with law enforcement.

A report from the Toronto Police Services highlights that individuals from Black, East/Southeast Asian, Middle Eastern, and Latino communities are overrepresented in use-of-force incidents compared to their presence in enforcement action populations. Despite efforts by the Toronto Police to promote fairness and equity, these findings suggest that racial bias may still persist in policing practices. However, a thorough, data-driven analysis of arrest patterns is necessary to determine the extent and nature of this bias. This paper addresses the gap by analyzing arrest data from 2020 and 2021 to examine how interactions between police and different ethnic groups are shaped by broader socio-economic factors such as poverty and unemployment. These factors, which disproportionately affect marginalized communities, may contribute to higher involvement in criminal activities, influencing arrest rates.

The analysis centers on arrest data, which provides an objective and quantifiable view of law enforcement activities, focusing on the correlation between arrests, ethnicity, and age. The data was processed and cleaned using R, followed by an analysis of the distribution of arrests across various age groups to identify potential trends and disparities in [@sec-data]. Age is considered a critical factor in understanding patterns of arrests, as it may reveal systemic biases in policing practices. The findings, discussed in [@sec-discussion], offer valuable insights into arrest patterns and their implications.

This study fills a crucial gap in the understanding of racial bias in policing in Toronto by providing an empirical assessment of arrest patterns across ethnic groups. It seeks to contribute to ongoing discussions of police reform by offering data-driven insights that can inform policy changes and promote community engagement. Furthermore, this analysis highlights the importance of transparency and accountability in policing, ensuring that law enforcement actions are equitable and just for all communities. engagement efforts.
